\LARGE \bf The Battle for Filter Supremacy: A Retrospective: The Movie: The Game: The Paper

### 2014-12-12 Report/project_report.tex
Minor notation change on the bias terms

The question, then, is how best to fuse measurements from these two sensor types to arrive at an accurate estimate of a vehicle's motion over time.
A complicating factor in this problem is the absence of a known map of landmarks from which the camera can generate measurements.
Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
What follows is a discussion of two common techniques, the Extended Kalman Filter (EKF) and the Sliding Window Filter (SWF), and a third hybrid approach, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.

@@

Although the recursive nature of EKF-SLAM allows it to operate online, the computational cost of the filter grows cubically with map size.
This behavior is due to the fact that the length of the state grows linearly with the number of landmarks, and the computational cost of inverting the state covariance matrix while computing the Kalman gain is cubic in the dimension of the state.
Consequently, the spatial extent over which EKF-SLAM can be used online is limited by the necessarily finite compute envelope available to it.

@@

An important advantage of the SWF is that its computational cost depends on the number of landmarks in the current window rather than the number of landmarks in the entire map.
By varying the spatial or temporal extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary spatial or temporal extent.

@@

We parametrize the IMU state at time $k$ as the 13-dimensional vector
\begin{align}
    \Vec{x}_{I,k} &= \bbm \Vec{q}_{IG,k} & \Vec{b}_{g,k} & \Vec{b}_{v,k} & \Vec{p}_{G,k}^{IG} \ebm^T
\end{align}
where $\Vec{q}_{IG,k}$ is the unit quaternion representing the rotation from the global frame $\mathcal{F}_G$ to the IMU frame $\mathcal{F}_I$, $\Vec{b}_{g,k}$ is the bias on the gyro measurements, $\Vec{b}_{v,k}$ is the bias on the velocity measurements, and $\Vec{p}_{G,k}^{IG}$ is the vector from the origin of $\mathcal{F}_G$ to the origin of $\mathcal{F}_I$ expressed in $\mathcal{F}_G$ (i.e., the position of the IMU in the global frame).

At time $k$, the state of the MSCKF consists of the current IMU state estimate, and estimates of $N$ past camera poses in which active feature tracks were visible:
\begin{align}
    \hat{\Vec{x}}_k &= \bbm \hat{\Vec{x}}_{I,k} & \hat{\Vec{q}}_{C_1 G} & \hat{\Vec{p}}_G^{C_1 G} & \hdots & \hat{\Vec{q}}_{C_N G} & \hat{\Vec{p}}_G^{C_N G} \ebm ^T
\end{align}.

\subsection{IMU State Estimate Propagation}
The evolution of the mean estimated IMU state $\hat{\Vec{x}}_I$ over time is described by the following differential equations:
\begin{align}
    \dot{\hat{\Vec{q}}}_{IG} &= \frac{1}{2}\boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right)\hat{\Vec{q}}_{IG} \\
    \dot{\hat{\Vec{b}}}_g  &= \Vec{0}_{3\times1} \\
    \dot{\hat{\Vec{b}}}_v &= \Vec{0}_{3\times1} \\
    \dot{\hat{\Vec{p}}}^G_I &= \hat{\Vec{C}}_{I G}^T\left(\hat{\Vec{v}}^G_I - \hat{\Vec{b}}_v \right)
\end{align}
where $\hat{\Vec{C}}_{I G}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG}$ and
\begin{align*}
    \boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right) &= \bbm -\hat{\boldsymbol{\omega}}^\times & \hat{\boldsymbol{\omega}} \\
                                                                -\hat{\boldsymbol{\omega}}^T & 0
                                                            \ebm 
\end{align*}
with
\begin{align*}
    \hat{\boldsymbol{\omega}} = \boldsymbol{\omega}_m - \hat{\Vec{b}}_g
\end{align*}
and
\begin{align*}
    \hat{\boldsymbol{\omega}}^\times &=   \bbm    0 & -\hat{\omega}_3 & \hat{\omega}_2 \\
                                                \hat{\omega}_3 & 0 & -\hat{\omega}_1 \\
                                                -\hat{\omega}_2 & \hat{\omega}_1 & 0
                                            \ebm .
\end{align*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{EXPERIMENTS} \label{sec:experiments}
We conducted a comparative study of the MSCKF and SWF algorithms using the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS).
The dataset consists of a rigidly attached stereo camera and IMU observing a set of 20 landmarks as they move along an arbitrary 3D path (Figure \ref{fig:sensorhead}).
The dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the landmark positions (Figure \ref{fig:trajectory_groundtruth}).
Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.

>>>>

The question, then, is how best to fuse measurements from these two sensor types to arrive at an accurate estimate of a vehicle's motion over time.
A complicating factor in the general form of this problem is the absence of a known map of landmarks from which the camera can generate measurements.
Any solution must therefore solve a Simultaneous Localization and Mapping (SLAM) problem, although the importance placed on the mapping component may vary from algorithm to algorithm.
What follows is a discussion of two common solutions, the Extended Kalman Filter (EKF) and the Sliding Window Filter (SWF), and a third hybrid solution, the Multi-State Constraint Kalman Filter (MSCKF), that combines the strengths of both.

@@

Although the recursive nature of EKF-SLAM allows it to operate online, the computational cost of the filter grows cubically with map size.
This behavior is due to the fact that the dimension of the state grows linearly with the number of landmarks, and the computational cost of inverting the state covariance matrix while computing the Kalman gain is cubic in the dimension of the state.
Consequently, the spatial extent over which EKF-SLAM can be used online is limited by the necessarily finite compute envelope available to it.

@@

An important advantage of the SWF is that its computational cost depends on the number of landmarks in the current window rather than the number of landmarks in the entire map.
By varying the spatial or temporal extent of the sliding window, the computational cost of the algorithm can be tailored to fit a given compute envelope, which makes the algorithm suitable for online operation over paths of arbitrary length.

@@

In our implementation, we parametrize the IMU state at time $k$ as the 13-dimensional vector
\begin{align}
    \Vec{x}_{I,k} &= \bbm \Vec{q}_{IG,k} & \Vec{b}_{\boldsymbol{\omega},k} & \Vec{b}_{\Vec{v},k} & \Vec{p}_{G,k}^{IG} \ebm^T
\end{align}
where $\Vec{q}_{IG,k}$ is the unit quaternion representing the rotation from the global frame $\mathcal{F}_G$ to the IMU frame $\mathcal{F}_I$, $\Vec{b}_{\boldsymbol{\omega},k}$ is the bias on the gyro measurements $\boldsymbol{\omega}_m$, $\Vec{b}_{\Vec{v},k}$ is the bias on the velocity measurements $\Vec{v}_m$, and $\Vec{p}_{G,k}^{IG}$ is the vector from the origin of $\mathcal{F}_G$ to the origin of $\mathcal{F}_I$ expressed in $\mathcal{F}_G$ (i.e., the position of the IMU in the global frame).

At time $k$, the full state of the MSCKF consists of the current IMU state estimate, and estimates of $N$ past camera poses in which active feature tracks were visible:
\begin{align}
    \hat{\Vec{x}}_k &= \bbm \hat{\Vec{x}}_{I,k} & \hat{\Vec{q}}_{C_1 G} & \hat{\Vec{p}}_G^{C_1 G} & \hdots & \hat{\Vec{q}}_{C_N G} & \hat{\Vec{p}}_G^{C_N G} \ebm ^T
\end{align}.

\subsection{IMU State Estimate Propagation}
The evolution of the mean estimated IMU state $\hat{\Vec{x}}_I$ over time is described by the following differential equations:
\begin{align}
    \dot{\hat{\Vec{q}}}_{IG} &= \frac{1}{2}\boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right)\hat{\Vec{q}}_{IG} \\
    \dot{\hat{\Vec{b}}}_{\boldsymbol{\omega}}  &= \Vec{0}_{3\times1} \\
    \dot{\hat{\Vec{b}}}_{\Vec{v}} &= \Vec{0}_{3\times1} \\
    \dot{\hat{\Vec{p}}}^{IG}_G &= \hat{\Vec{C}}_{I G}^T\left(\Vec{v}_m - \hat{\Vec{b}}_{\Vec{v}} \right)
\end{align}
where $\hat{\Vec{C}}_{I G}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG}$ and
\begin{align*}
    \boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right) &= \bbm -\hat{\boldsymbol{\omega}}^\times & \hat{\boldsymbol{\omega}} \\
                                                                -\hat{\boldsymbol{\omega}}^T & 0
                                                            \ebm 
\end{align*}
with
\begin{align*}
    \hat{\boldsymbol{\omega}} = \boldsymbol{\omega}_m - \hat{\Vec{b}}_{\boldsymbol{\omega}}
\end{align*}
and
\begin{align*}
    \hat{\boldsymbol{\omega}}^\times &=   \bbm    0 & -\hat{\omega}_3 & \hat{\omega}_2 \\
                                                \hat{\omega}_3 & 0 & -\hat{\omega}_1 \\
                                                -\hat{\omega}_2 & \hat{\omega}_1 & 0
                                            \ebm .
\end{align*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{EXPERIMENTS} \label{sec:experiments}
We conducted a comparative study of the MSCKF and SWF algorithms using the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS).
The dataset consists of a rigidly attached stereo camera and IMU (Figure \ref{fig:sensorhead}) observing a set of 20 landmarks while moving along an arbitrary 3D path.
The dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the landmark positions (Figure \ref{fig:trajectory_groundtruth}).
Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.
